---
title: "Mutation Burden vs Age Testicular normal samples Signature Analyses"
---

#Load Libraries
```{r}
library('biomaRt')
library(patchwork)
library(rtracklayer)
library(stringr)
library(stats)
library(devtools)
library("cowplot")
library("dplyr")
library("ggplot2")
library("ggpubr")
library("gridExtra")
library("lme4")
library("multcomp")
library(MutationalPatterns)
library("nlme")
library("plyr")
library("randomcoloR")
library("RColorBrewer")
library("reshape2")
library("tidyverse")
library(BSgenome)
library("TCGAbiolinks")
library(org.Hs.eg.db)
library(Organism.dplyr)
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library("TxDb.Hsapiens.UCSC.hg19.knownGene")
library(ref_genome, character.only = TRUE)
options(scipen = 999)
theme_set(theme_cowplot())






```

# Full dataset
```{r}
vcf_files <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Inputs/All", pattern = "\\_complete_final_retained_4.goodSNPS_appended.manualFiltered.vcf$", full.names = TRUE)

sample_names <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Inputs/All", pattern = "\\_complete_final_retained_4.goodSNPS_appended.manualFiltered.vcf$")

sample_names <- str_remove(sample_names, "_complete_final_retained_4.goodSNPS_appended.manualFiltered.vcf")

surveyed_files <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Surveyed_regions", pattern = "\\.callable.bed$", full.names = TRUE)
surveyed_list <- vector(mode = "list")

for(i in sample_names){
  surveyed <- import(paste("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Surveyed_regions/",i,".callable.bed", sep=""))
  ## Import the file using rtracklayer and use the UCSC naming standard
  seqlevelsStyle(surveyed) <- "UCSC"
  surveyed_list[[i]] <- surveyed
}

length(vcf_files[234])
head(surveyed_list)
PID <- substr(sample_names,0,7)
PIDTissue <- substr(sample_names,0,8)
vcfs <- read_vcfs_as_granges(vcf_files, sample_names, ref_genome, check_alleles=T)
#15 position(s) with indels and/or multiple alternative alleles are excluded from Trio_DNMs.
summary(vcfs)
muts = mutations_from_vcf(vcfs[[1]])
types = mut_type(vcfs[[1]])
context = mut_context(vcfs[[1]], ref_genome)
type_context = type_context(vcfs[[1]], ref_genome)
type_occurrences <- mut_type_occurrences(vcfs, ref_genome)

Tumour_Normal_tissue_info <- read.table("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/metadata/Tumour_Normal_Unmatched-WGS_analysis.Normal-Colon-Benign.txt", header=T, check.names=FALSE, sep="\t",quote = "")

Tissue <- merge.data.frame(as.data.frame(sample_names),Tumour_Normal_tissue_info, by.x="sample_names", by.y="Tumour")
Tissue <- Tissue[,"Tumour_Tissue_Type"]
sample_info <- read.table("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/metadata/Sample_annotation.txt", header=T, check.names=FALSE, sep="\t",quote = "")
unique(sample_info$Annotation)

sample_info <- sample_info[sample_info$Annotation %in% c("Normal", "Tumour_background-Normal","Not_normal(benign)", "Colon", "Trio_DNMs"),]
sample_info <- sample_info[order(sample_info$Age),]
rownames(sample_info) <- NULL
sample_info$Sample <- factor(sample_info$Sample, levels=sample_info$Sample)
sort_order <- unique(str_remove_all(sample_info$Sample, "[bcf]"))
#sort_order <- factor(c("PD43725","PD42036","PD40744","PD42034","PD42563","PD42569","PD40745","PD42564", "PD42565","PD42566","PD42038","PD43727","PD40746","PD42568", "PD43724","PD43726"), levels=c("PD43725","PD42036","PD40744","PD42034","PD42563","PD42569","PD40745","PD42564", "PD42565","PD42566","PD42038","PD43727","PD40746","PD42568", "PD43724","PD43726"))
sort_order <- factor(sort_order, levels=sort_order)


revalue(PID, c("PD40745" = "PD40745/PD42564")) -> PID
revalue(PID, c("PD42564" = "PD40745/PD42564")) -> PID
revalue(PID, c("PD40746" = "PD40746/PD42568")) -> PID
revalue(PID, c("PD42568" = "PD40746/PD42568")) -> PID

revalue(PIDTissue, c("PD40745c" = "PD40745c/PD42564c")) -> PIDTissue
revalue(PIDTissue, c("PD42564c" = "PD40745c/PD42564c")) -> PIDTissue
revalue(PIDTissue, c("PD40746c" = "PD40746c/PD42568c")) -> PIDTissue
revalue(PIDTissue, c("PD42568c" = "PD40746c/PD42568c")) -> PIDTissue

revalue(PIDTissue, c("PD40745b" = "PD40745b/PD42564b")) -> PIDTissue
revalue(PIDTissue, c("PD42564b" = "PD40745b/PD42564b")) -> PIDTissue
revalue(PIDTissue, c("PD40746b" = "PD40746b/PD42568b")) -> PIDTissue
revalue(PIDTissue, c("PD42568b" = "PD40746b/PD42568b")) -> PIDTissue

length(names(vcfs))
length(PID)
length(Tissue)
length(sample_names)
unique(sort_order)

PID <- as.character(factor(PID, levels=sort_order))

sort_order_PIDTissue <- unique(sample_info$Sample)
sort_order_PIDTissue <- factor(sort_order_PIDTissue, levels=sort_order_PIDTissue)

PIDTissue <- as.character(factor(PIDTissue, levels=sort_order_PIDTissue))

#Merging tumour background normal and normal into single set
Tissue_Normal_TBN_merged  <- as.character(Tissue) %>% 
  str_replace("Tumour_background-Normal","Normal") %>% 
  str_replace("Normal","Testes") %>% 
  str_replace("Not_normal(benign)","Testes(benign)")
```

# Trio dataset
```{r}
#Generate trio vcf files separately
trio_vcf_files <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Trio_data/per_sample", pattern = "\\.vcf$", full.names = TRUE)

trio_sample_names <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Trio_data/per_sample", pattern = "\\.vcf$")

trio_sample_names <- str_remove(trio_sample_names, ".vcf")

trio_surveyed_file <- "/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Surveyed_regions/Trio_DNMs.callable.bed"
trio_surveyed <- import(trio_surveyed_file)
seqlevelsStyle(trio_surveyed) <- "UCSC"
trio_surveyed_list <- rep(list(trio_surveyed), 3450)


trio_PID <- trio_sample_names
trio_PIDTissue <- as.list(c(rep("asd", 1902), rep("decode",1548)))

trio_vcfs <- read_vcfs_as_granges(trio_vcf_files, trio_sample_names, ref_genome, check_alleles=T)
#15 position(s) with indels and/or multiple alternative alleles are excluded from Trio_DNMs.
summary(trio_vcfs)
trio_muts = mutations_from_vcf(trio_vcfs[[1]])
trio_types = mut_type(trio_vcfs[[1]])
trio_context = mut_context(trio_vcfs[[1]], ref_genome)
trio_type_context = type_context(trio_vcfs[[1]], ref_genome)
trio_type_occurrences <- mut_type_occurrences(trio_vcfs, ref_genome)


trio_Tissue <- as.list(rep("Trio_DNMs",3450))

load("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/mutational_pattern_LCMsamples.rds")




```


# 6 Mutation Type plots
```{r}
#Merge trios data with LCM
merged_type_occurrences <- rbind(type_occurrences, trio_type_occurrences)

merged_PID <- c(PID, trio_PID)
merged_PIDTissue <- c(PIDTissue, trio_PIDTissue)
merged_Tissue <- as.character(c(Tissue_Normal_TBN_merged, trio_Tissue))

png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/MutPat-ByPIDTissue-6mutTypes_withTrioData.png", res=300, units="in", width = 12, height=7)
plot_spectrum(merged_type_occurrences, by= merged_Tissue, CT = TRUE, legend = TRUE)
dev.off()

```

# 96 Trinucleotide mutational spectrum plots
```{r}
mutational_catalog_colors <- c("#2EBAED", "#000000", "#DE1C14","#D4D2D2", "#ADCC54", "#F0D0CE")
names(mutational_catalog_colors) <- c("C>A", "C>G", "C>T", "T>A", "T>C", "T>G")

mut_mat <- mut_matrix(vcf_list = vcfs, ref_genome = ref_genome)
colnames(mut_mat)

save(vcfs,PID,PIDTissue, Tissue,Tissue_Normal_TBN_merged,sample_names,mutational_catalog_colors, surveyed_list, file = "/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/mutational_pattern_allsamples.rds")
load("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/mutational_pattern_allsamples.rds")


# Per sample
for(i in colnames(mut_mat)){
  mut_mat_per_Sample <- as.data.frame(mut_mat[,i])
  colnames(mut_mat_per_Sample) <- i
  mat <- as.matrix(mut_mat_per_Sample)
  ymax <- max(mat/sum(mat))
  p <- plot_96_profile(mut_mat_per_Sample, condensed = TRUE, ymax=ymax+0.02)
  outfile <- paste("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Per_sample_Mutational_context_plots/muta_spectrum_plot",i,".png",sep="")
  png(file=outfile, res=300, units="in", height = 3, width=9)
  print(p)
  dev.off()
}


# Per Tissue ("Testes", "Colon", "Testes(Benign)", Trio_DNMs)  with 95% confidence intervals
for(i in unique(Tissue_Normal_TBN_merged)){
  rownumbers <- which(Tissue_Normal_TBN_merged == i)
  mut_mat_Tmp <- mut_mat[,rownumbers]
  mut_mat_Tmp_mat <- as.matrix(mut_mat_Tmp)
  mut_mat_Tmp_mat <- mut_mat_Tmp_mat/colSums(mut_mat_Tmp_mat)
  mut_mat_Tmp <- as.data.frame(mut_mat_Tmp_mat)
  mut_mat_plot <- data.frame(matrix(ncol = 2, nrow = length(rownames(mut_mat_Tmp))))
  colnames(mut_mat_plot) <- c("mean", "ci95")
  rownames(mut_mat_plot) <- rownames(mut_mat_Tmp)
  for(j in rownames(mut_mat_Tmp)){
     info <- summarySE(t(mut_mat_Tmp),measurevar=j,na.rm = FALSE, conf.interval = 0.95)
     mut_mat_plot[j,"mean"] <- info[,j]
     mut_mat_plot[j,"ci95"] <- info[,"ci"]
  }
  mut_mat_plot[,"Context"] <- paste(substr(rownames(mut_mat_plot),1,1),substr(rownames(mut_mat_plot),3,3),substr(rownames(mut_mat_plot),7,7), sep="")
  mut_mat_plot[,"Type"] <- substr(rownames(mut_mat_plot),3,5)
  p <- ggplot(mut_mat_plot, aes(x=Context, y=mean, fill=Type)) + geom_bar(stat="identity", color="black", position=position_dodge()) + geom_errorbar(aes(ymin=mean-ci95, ymax=mean+ci95), width=.2, position=position_dodge(.9)) + facet_grid(.~Type, scales="free", space = "free") + scale_fill_manual(values=mutational_catalog_colors)+theme_pubr()+theme(legend.position="right", legend.direction="vertical", panel.spacing = unit(0, "lines"), axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=5))+ylab("Relative Contribution") + xlab("")
  outfile <- paste("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Per_TissueType_Mutational_context_plots/muta_spectrum_plot.",i,".png",sep="")
  png(file=outfile, res=300, units="in", height = 3, width=9)
  print(p)
  dev.off()
}

# Per PID  with 95% confidence intervals
for(i in unique(PID)){
  rownumbers <- which(PID == i)
  mut_mat_Tmp <- mut_mat[,rownumbers]
  mut_mat_Tmp_mat <- as.matrix(mut_mat_Tmp)
  mut_mat_Tmp_mat <- mut_mat_Tmp_mat/colSums(mut_mat_Tmp_mat)
  mut_mat_Tmp <- as.data.frame(mut_mat_Tmp_mat)
  mut_mat_plot <- data.frame(matrix(ncol = 2, nrow = length(rownames(mut_mat_Tmp))))
  colnames(mut_mat_plot) <- c("mean", "ci95")
  rownames(mut_mat_plot) <- rownames(mut_mat_Tmp)
  for(j in rownames(mut_mat_Tmp)){
     info <- summarySE(t(mut_mat_Tmp),measurevar=j,na.rm = FALSE, conf.interval = 0.95)
     mut_mat_plot[j,"mean"] <- info[,j]
     mut_mat_plot[j,"ci95"] <- info[,"ci"]
  }
  mut_mat_plot[,"Context"] <- paste(substr(rownames(mut_mat_plot),1,1),substr(rownames(mut_mat_plot),3,3),substr(rownames(mut_mat_plot),7,7), sep="")
  mut_mat_plot[,"Type"] <- substr(rownames(mut_mat_plot),3,5)
  p <- ggplot(mut_mat_plot, aes(x=Context, y=mean, fill=Type)) + geom_bar(stat="identity", color="black", position=position_dodge()) + geom_errorbar(aes(ymin=mean-ci95, ymax=mean+ci95), width=.2, position=position_dodge(.9)) + facet_grid(.~Type, scales="free", space = "free") + scale_fill_manual(values=mutational_catalog_colors)+theme_pubr()+theme(legend.position="right", legend.direction="vertical", panel.spacing = unit(0, "lines"), axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=5))+ylab("Relative Contribution") + xlab("")
  i <- str_replace(i, "/","_")
  outfile <- paste("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Per_PID_Mutational_context_plots/muta_spectrum_plot.",i,".png",sep="")
  png(file=outfile, res=300, units="in", height = 3, width=9)
  print(p)
  dev.off()
}

# Per PID sparated by tissue with 95% confidence intervals
for(i in unique(PIDTissue)){
  rownumbers <- which(PIDTissue == i)
  mut_mat_Tmp <- mut_mat[,rownumbers]
  mut_mat_Tmp_mat <- as.matrix(mut_mat_Tmp)
  mut_mat_Tmp_mat <- mut_mat_Tmp_mat/colSums(mut_mat_Tmp_mat)
  mut_mat_Tmp <- as.data.frame(mut_mat_Tmp_mat)
  mut_mat_plot <- data.frame(matrix(ncol = 2, nrow = length(rownames(mut_mat_Tmp))))
  colnames(mut_mat_plot) <- c("mean", "ci95")
  rownames(mut_mat_plot) <- rownames(mut_mat_Tmp)
  for(j in rownames(mut_mat_Tmp)){
     info <- summarySE(t(mut_mat_Tmp),measurevar=j,na.rm = FALSE, conf.interval = 0.95)
     mut_mat_plot[j,"mean"] <- info[,j]
     mut_mat_plot[j,"ci95"] <- info[,"ci"]
  }
  mut_mat_plot[,"Context"] <- paste(substr(rownames(mut_mat_plot),1,1),substr(rownames(mut_mat_plot),3,3),substr(rownames(mut_mat_plot),7,7), sep="")
  mut_mat_plot[,"Type"] <- substr(rownames(mut_mat_plot),3,5)
  p <- ggplot(mut_mat_plot, aes(x=Context, y=mean, fill=Type)) + geom_bar(stat="identity", color="black", position=position_dodge()) + geom_errorbar(aes(ymin=mean-ci95, ymax=mean+ci95), width=.2, position=position_dodge(.9)) + facet_grid(.~Type, scales="free", space = "free") + scale_fill_manual(values=mutational_catalog_colors)+theme_pubr()+theme(legend.position="right", legend.direction="vertical", panel.spacing = unit(0, "lines"), axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=5))+ylab("Relative Contribution") + xlab("")
  i <- str_replace(i, "/","_")
  outfile <- paste("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Per_PIDTissue_Mutational_context_plots/muta_spectrum_plot.",i,".png",sep="")
  png(file=outfile, res=300, units="in", height = 3, width=9)
  print(p)
  dev.off()
}

# Per PID with STDDEV
for(i in unique(PID)){
  rownumbers <- which(PID == i)
  mut_mat_Tmp <- mut_mat[,rownumbers]
  mut_mat_Tmp_mat <- as.matrix(mut_mat_Tmp)
  mut_mat_Tmp_mat <- mut_mat_Tmp_mat/colSums(mut_mat_Tmp_mat)
  mut_mat_Tmp <- as.data.frame(mut_mat_Tmp_mat)
  mut_mat_plot <- data.frame(matrix(ncol = 2, nrow = length(rownames(mut_mat_Tmp))))
  colnames(mut_mat_plot) <- c("mean", "sd")
  rownames(mut_mat_plot) <- rownames(mut_mat_Tmp)
  for(j in rownames(mut_mat_Tmp)){
     info <- summarySE(t(mut_mat_Tmp),measurevar=j,na.rm = FALSE, conf.interval = 0.95)
     mut_mat_plot[j,"mean"] <- info[,j]
     mut_mat_plot[j,"sd"] <- info[,"sd"]
  }
  mut_mat_plot[,"Context"] <- paste(substr(rownames(mut_mat_plot),1,1),substr(rownames(mut_mat_plot),3,3),substr(rownames(mut_mat_plot),7,7), sep="")
  mut_mat_plot[,"Type"] <- substr(rownames(mut_mat_plot),3,5)
  p <- ggplot(mut_mat_plot, aes(x=Context, y=mean, fill=Type)) + geom_bar(stat="identity", color="black", position=position_dodge()) + geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2, position=position_dodge(.9)) + facet_grid(.~Type, scales="free", space = "free") + scale_fill_manual(values=mutational_catalog_colors)+theme_pubr()+theme(legend.position="right", legend.direction="vertical", panel.spacing = unit(0, "lines"), axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=5))+ylab("Relative Contribution") + xlab("")
  i <- str_replace(i, "/","_")
  outfile <- paste("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Per_PID_Mutational_context_plots/muta_spectrum_plot.",i,"withSTDev.png",sep="")
  png(file=outfile, res=300, units="in", height = 3, width=9)
  print(p)
  dev.off()
}

# Per PID sparated by tissue with STDDEV
for(i in unique(PIDTissue)){
  rownumbers <- which(PIDTissue == i)
  mut_mat_Tmp <- mut_mat[,rownumbers]
  mut_mat_Tmp_mat <- as.matrix(mut_mat_Tmp)
  mut_mat_Tmp_mat <- mut_mat_Tmp_mat/colSums(mut_mat_Tmp_mat)
  mut_mat_Tmp <- as.data.frame(mut_mat_Tmp_mat)
  mut_mat_plot <- data.frame(matrix(ncol = 2, nrow = length(rownames(mut_mat_Tmp))))
  colnames(mut_mat_plot) <- c("mean", "sd")
  rownames(mut_mat_plot) <- rownames(mut_mat_Tmp)
  for(j in rownames(mut_mat_Tmp)){
     info <- summarySE(t(mut_mat_Tmp),measurevar=j,na.rm = FALSE, conf.interval = 0.95)
     mut_mat_plot[j,"mean"] <- info[,j]
     mut_mat_plot[j,"sd"] <- info[,"sd"]
  }
  mut_mat_plot[,"Context"] <- paste(substr(rownames(mut_mat_plot),1,1),substr(rownames(mut_mat_plot),3,3),substr(rownames(mut_mat_plot),7,7), sep="")
  mut_mat_plot[,"Type"] <- substr(rownames(mut_mat_plot),3,5)
  p <- ggplot(mut_mat_plot, aes(x=Context, y=mean, fill=Type)) + geom_bar(stat="identity", color="black", position=position_dodge()) + geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2, position=position_dodge(.9)) + facet_grid(.~Type, scales="free", space = "free") + scale_fill_manual(values=mutational_catalog_colors)+theme_pubr()+theme(legend.position="right", legend.direction="vertical", panel.spacing = unit(0, "lines"), axis.text.x = element_text(angle=90, vjust=0.5, hjust=0, size=5))+ylab("Relative Contribution") + xlab("")
  i <- str_replace(i, "/","_")
  outfile <- paste("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Per_PIDTissue_Mutational_context_plots/muta_spectrum_plot.",i,"withSTDev.png",sep="")
  png(file=outfile, res=300, units="in", height = 3, width=9)
  print(p)
  dev.off()
}

```


##Transcriptional strand bias analysis
```{r}
ymax=max()
genes_hg19 <- genes(TxDb.Hsapiens.UCSC.hg19.knownGene)
mut_mat_s <- mut_matrix_stranded(vcfs, ref_genome, genes_hg19)

#By Tissue Type ("Testes", "Colon", "Testes(Benign)")
strand_counts <- strand_occurrences(mut_mat_s, by=Tissue_Normal_TBN_merged)
#Poisson test for strand asymmetry significance
strand_bias <- strand_bias_test(strand_counts)
#Plot the mutation spectrum with strand distinction:
ps1 <- plot_strand(strand_counts, mode = "relative")
#effect size (log2(untranscribed/transcribed) of the strand bias
ps2 <- plot_strand_bias(strand_bias)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Transcriptional_strand_bias_plot_ByTissue.png", res=300, units="in", width = 9, height=6)
ps1 / ps2
dev.off()

#Per PID seperated by Tissue
strand_counts <- strand_occurrences(mut_mat_s, by=PIDTissue)
#Poisson test for strand asymmetry significance
strand_bias <- strand_bias_test(strand_counts)
#Plot the mutation spectrum with strand distinction:
ps1 <- plot_strand(strand_counts, mode = "relative")
#effect size (log2(untranscribed/transcribed) of the strand bias
ps2 <- plot_strand_bias(strand_bias)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Transcriptional_strand_bias_plot_ByPIDTissue.png", res=300, units="in", width = 16, height=6)
ps1 / ps2
dev.off()

#For Testes Samples per PID
strand_counts <- strand_occurrences(mut_mat_s, by=PIDTissue)
rownumbers <- which(Tissue_Normal_TBN_merged == "Testes")
testes_names <- unique(PIDTissue[rownumbers])
strand_counts <- strand_counts[which(strand_counts$group %in% testes_names),]
#Poisson test for strand asymmetry significance
strand_bias <- strand_bias_test(strand_counts)
#Plot the mutation spectrum with strand distinction:
ps1 <- plot_strand(strand_counts, mode = "relative")
#effect size (log2(untranscribed/transcribed) of the strand bias
ps2 <- plot_strand_bias(strand_bias)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Transcriptional_strand_bias_plot_ByPIDTestes.png", res=300, units="in", width = 16, height=6)
ps1 / ps2
dev.off()

#For Colon Samples per PID
strand_counts <- strand_occurrences(mut_mat_s, by=PIDTissue)
rownumbers <- which(Tissue_Normal_TBN_merged == "Colon")
Colon_names <- unique(PIDTissue[rownumbers])
strand_counts <- strand_counts[which(strand_counts$group %in% Colon_names),]
#Poisson test for strand asymmetry significance
strand_bias <- strand_bias_test(strand_counts)
#Plot the mutation spectrum with strand distinction:
ps1 <- plot_strand(strand_counts, mode = "relative")
#effect size (log2(untranscribed/transcribed) of the strand bias
ps2 <- plot_strand_bias(strand_bias)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/Transcriptional_strand_bias_plot_ByPIDColon.png", res=300, units="in", width = 8, height=6)
ps1 / ps2
dev.off()


```

# Looking at Genic and Non Genic regions
```{r}
# reduce

src <- src_organism("TxDb.Hsapiens.UCSC.hg19.knownGene")

txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene 
head(seqlevels(txdb))
columns(txdb)
txdb$GENEID
coding <- cds(txdb)
coding_df <- data.frame(Chr=seqnames(coding), Start=start(coding)-1, End=end(coding), Strand=strand(coding))
write.table(coding_df,file="/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/External_Data-Code/Tissue-specific_mutation_accumulation-paper/Data/genomic_regions/Homo_sapiens.GRCh37.75_coding.bed", sep="\t", row.names = FALSE, quote = FALSE)

exons_hg19 <- exons(txdb)
exons_hg19_df <- data.frame(Chr=seqnames(exons_hg19), Start=start(exons_hg19)-1, End=end(exons_hg19), Strand=strand(exons_hg19))
write.table(exons_hg19_df,file="/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/External_Data-Code/Tissue-specific_mutation_accumulation-paper/Data/genomic_regions/Homo_sapiens.GRCh37.75_exons.bed", sep="\t", row.names = FALSE, quote = FALSE)

#cdsBy(txdb, by = c("tx", "gene"))


# Get genes and non-genic genomic regions
hg19_Granges = GRanges(seqnames = seqnames(get(ref_genome))[1:24], IRanges(start=rep(1,24), end=seqlengths(get(ref_genome))[1:24]))
hg19_nongenic = setdiff(hg19_Granges, genes_hg19, ignore.strand=T)
# Combine all genomic regions (GRanges objects) in a named list:
regions = list(genes_hg19, hg19_nongenic)
names(regions) = c("Genes", "Non-genic")
distr <- genomic_distribution(vcfs, surveyed_list, regions)


distr_test <- enrichment_depletion_test(distr, by=Tissue_Normal_TBN_merged)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/Genic_vs_NonGenic-ByTissue.png", res=300, units="in", width = 5, height=5)
plot_enrichment_depletion(distr_test)
dev.off()

distr_test <- enrichment_depletion_test(distr, by=PIDTissue)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/Genic_vs_NonGenic-ByPIDTissue.png", res=300, units="in", width = 8, height=8)
plot_enrichment_depletion(distr_test)
dev.off()


```

# Plot Genomic regions."Promoter", "Promoter flanking", "CTCF"
```{r}

CTCF_g <- readRDS(system.file("states/CTCF_g_data.rds", package="MutationalPatterns"))
promoter_g <- readRDS(system.file("states/promoter_g_data.rds", package="MutationalPatterns"))
flanking_g <- readRDS(system.file("states/promoter_flanking_g_data.rds", package="MutationalPatterns"))


regions <- GRangesList(promoter_g, flanking_g, CTCF_g)
names(regions) <- c("Promoter", "Promoter flanking", "CTCF")
seqlevelsStyle(regions) <- "UCSC"

distr <- genomic_distribution(vcfs, surveyed_list, regions)

distr_test <- enrichment_depletion_test(distr,by=Tissue_Normal_TBN_merged)
plot_enrichment_depletion(distr_test)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/CTCF_Promoter_Flanking-ByTissue.png", res=300, units="in", width = 6, height=5)
plot_enrichment_depletion(distr_test)
dev.off()

distr_test <- enrichment_depletion_test(distr,by=PIDTissue)
plot_enrichment_depletion(distr_test)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/CTCF_Promoter_Flanking-ByPIDTissue.png", res=300, units="in", width = 8, height=10)
plot_enrichment_depletion(distr_test)
dev.off()
```



# Replication Timming
## Method from Harald(TensorSignatures)
```{r}
#' Replication times
repTimeGm12878 <- import("http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwRepliSeq//wgEncodeUwRepliSeqGm12878WaveSignalRep1.bigWig", format="BigWig")
repTimeK569 <- import("http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwRepliSeq/wgEncodeUwRepliSeqK562WaveSignalRep1.bigWig", format="BigWig")
repTimeHela <- import("http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwRepliSeq/wgEncodeUwRepliSeqHelas3WaveSignalRep1.bigWig", format="BigWig")
repTimeHuvec <- import("http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwRepliSeq/wgEncodeUwRepliSeqHuvecWaveSignalRep1.bigWig", format="BigWig")
repTimeHepg2 <- import("http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwRepliSeq/wgEncodeUwRepliSeqHepg2WaveSignalRep1.bigWig", format="BigWig")

repTime <- granges(repTimeGm12878)
mcols(repTime) <- DataFrame(mg12878=repTimeGm12878$score, hela=repTimeHela$score, k562=repTimeK569$score, huvec=repTimeHuvec$score, hepg2=repTimeHepg2$score)
ndiff <- function(x) (c(NA,x[-length(x)]) - c(x[-1],NA))
t <- sapply(mcols(repTime), ndiff)
m <- rowMeans(t)
s <- rowMeans(sqrt((t-m)^2))
table(abs(m)-2*s > 0)

repStrand <- granges(repTime)
strand(repStrand)[which(m<0)] <- "+" 
strand(repStrand)[which(m>0)] <- "-" 
strand(repStrand)[which(abs(m)-2*s < 0)] <- "*" 
seqlevels(repStrand) <- sub("chr","", seqlevels(repStrand))

save(repStrand, file="/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/repStrand.RData")

load("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Unmatched_WGS/Signature_Analyses/repStrand.RData")
getRepStrand <- function(vcf, repStrand){
f <- findOverlaps(vcf, repStrand)
s <- as.factor(strand(repStrand))[selectHits(f, select="first")]
return(s)
}
seqlevelsStyle(repStrand) <- "UCSC"
Leading_strand <- repStrand[which(strand(repStrand) == "+")]
Lagging_strand <- repStrand[which(strand(repStrand) == "-")]
Unassigned_strand <- repStrand[which(strand(repStrand) == "*")]

replicationDirection_list <- GRangesList(Leading_strand, Lagging_strand, Unassigned_strand)
names(replicationDirection_list) <- c("Leading", "Lagging", "Unassigned")

gen_dist = genomic_distribution(vcfs, surveyed_list, replicationDirection_list)

distr_test <- enrichment_depletion_test(gen_dist,by=Tissue_Normal_TBN_merged)
plot_enrichment_depletion(distr_test)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/Replication_direction-ByTissue.png", res=300, units="in", width = 6, height=5)
plot_enrichment_depletion(distr_test)
dev.off()

distr_test <- enrichment_depletion_test(gen_dist,by=PIDTissue)
plot_enrichment_depletion(distr_test)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/Replication_direction-ByPIDTissue.png", res=300, units="in", width = 10, height=5)
plot_enrichment_depletion(distr_test)
dev.off()



```

## Method from https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5536223/pdf/emss-73438.pdf
```{r}
# regions bed files
regions_files = list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/External_Data-Code/Tissue-specific_mutation_accumulation-paper/Data/genomic_regions", full.names = T)
print(regions_files)
# region bed files
regions_list=list()
for(i in regions_files){
  print(i)
  region <- import(i)
  ## Import the file using rtracklayer and use the UCSC naming standard
  seqlevelsStyle(region) <- "UCSC"
  regions_list[[i]] <- region
}

names(regions_list) = c("H3k27ac", "H3K9me3", "Early", "Intermediate", "Late", "Coding")

# Find overlaps between mutations and genomic regions
# dit draaien zonder subset
gen_dist = genomic_distribution(vcfs, surveyed_list, regions_list)

enr_depl = enrichment_depletion_test(gen_dist, by = Tissue_Normal_TBN_merged)
enr_depl$by = factor(enr_depl$by, levels = c("Colon", "Testes", "Testes(benign)","Trio_DNMs"))
enr_depl$region = factor(enr_depl$region, levels = c("Coding", "Early", "Intermediate", "Late",  "H3k27ac", "H3K9me3"))
plot_enrichment_depletion(enr_depl)


png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/enrichment_depletion-byTissue.png", width=8, height=6, res=300, units="in")
plot_enrichment_depletion(enr_depl)
dev.off()

enr_depl = enrichment_depletion_test(gen_dist, by = PIDTissue)
enr_depl$region = factor(enr_depl$region, levels = c("Coding", "Early", "Intermediate", "Late",  "H3k27ac", "H3K9me3"))

png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Final_Analyses/Enrichment_Analyses/Genomic_regions/enrichment_depletion-byPIDTissue.png", width=16, height=6, res=300, units="in")
plot_enrichment_depletion(enr_depl)
dev.off()

```
## Rainfall plots
```{r}

chromosomes <- seqnames(get(ref_genome))[1:22]

rownumbers <- which(Tissue_Normal_TBN_merged == "Testes")
sort_order_normal_tmp <- unique(PID[rownumbers])

revalue(sort_order_normal_tmp, c("PD40745/PD42564" ="PD40745")) -> sort_order_normal_tmp
revalue(sort_order_normal_tmp, c("PD40746/PD42568" ="PD40746")) -> sort_order_normal_tmp

p_rainfall <- plot_rainfall(vcfs$PD40744, title = "PD40744", chromosomes = chromosomes, cex = 1.5, ylim = 1e+09)
for(name in sort_order_normal_tmp){
  if(name!="PD40744" ){
    p_rainfall <- p_rainfall + plot_rainfall(vcfs[[name]], title = name, chromosomes = chromosomes, cex = 1.5, ylim = 1e+09)
  }
}
p_rainfall

png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/Plots/Rainfall_plot_normals_only-maftools.png", res=300, units="in", width = 16, height=8)
p_rainfall
dev.off()
```

# Early acquired variants vs Late acquired Variants
##Early
```{r}
vcf_files_early <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/variants_filtering-2019_12_05/Good_SNPS_Normal-Testes_only/VCF_per_ID", pattern = "\\.early.vcf$", full.names = TRUE)
sample_names <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/variants_filtering-2019_12_05/Good_SNPS_Normal-Testes_only/VCF_per_ID", pattern = "\\.early.vcf$")
sample_names <- str_remove(sample_names, ".early.vcf")
vcfs_early <- read_vcfs_as_granges(vcf_files_early, sample_names, ref_genome)
summary(vcfs_early)
muts_early = mutations_from_vcf(vcfs_early[[1]])
types_early = mut_type(vcfs_early[[1]])
context_early = mut_context(vcfs_early[[1]], ref_genome)
type_context_early = type_context(vcfs_early[[1]], ref_genome)
type_occurrences_early <- mut_type_occurrences(vcfs_early, ref_genome)

sample_info <- read.table("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/analysis/Sample_annotation.txt", header=T, check.names=FALSE, sep="\t",quote = "")
unique(sample_info$Annotation)

sample_info <- sample_info[sample_info$Annotation %in% c("Normal", "Tumour_background-Normal","Not_normal(benign)"),]
sample_info <- sample_info[order(sample_info$Age),]
rownames(sample_info) <- NULL
sample_info$Sample <- factor(sample_info$Sample, levels=sample_info$Sample)
sort_order <- str_remove_all(sample_info$Sample, "[bcf]")
#sort_order <- factor(c("PD43725","PD42036","PD40744","PD42034","PD42563","PD42569","PD40745","PD42564", "PD42565","PD42566","PD42038","PD43727","PD40746","PD42568", "PD43724","PD43726"), levels=c("PD43725","PD42036","PD40744","PD42034","PD42563","PD42569","PD40745","PD42564", "PD42565","PD42566","PD42038","PD43727","PD40746","PD42568", "PD43724","PD43726"))
sort_order <- factor(sort_order, levels=sort_order)


p3 <- plot_spectrum(type_occurrences_early, CT = TRUE, legend = FALSE)
p4 <- plot_spectrum(type_occurrences_early, by = sort_order, CT = TRUE, legend = TRUE)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/Plots/Mutation_type_plot-maftools.png", res=300, units="in", width = 12, height=7)
grid.arrange(p3, p4, ncol=2, widths=c(1,3))
dev.off()
```


##Late
```{r}
vcf_files_late <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/variants_filtering-2019_12_05/Good_SNPS_Normal-Testes_only/VCF_per_ID", pattern = "\\.late.vcf$", full.names = TRUE)
sample_names <- list.files("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/variants_filtering-2019_12_05/Good_SNPS_Normal-Testes_only/VCF_per_ID", pattern = "\\.late.vcf$")
sample_names <- str_remove(sample_names, ".late.vcf")
vcfs_late <- read_vcfs_as_granges(vcf_files_late, sample_names, ref_genome)
summary(vcfs_late)
muts_late = mutations_from_vcf(vcfs_late[[1]])
types_late = mut_type(vcfs_late[[1]])
context_late = mut_context(vcfs_late[[1]], ref_genome)
type_context_late = type_context(vcfs_late[[1]], ref_genome)
type_occurrences_late <- mut_type_occurrences(vcfs_late, ref_genome)

sample_info <- read.table("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/analysis/Sample_annotation.txt", header=T, check.names=FALSE, sep="\t",quote = "")
unique(sample_info$Annotation)

sample_info <- sample_info[sample_info$Annotation %in% c("Normal", "Tumour_background-Normal","Not_normal(benign)"),]
sample_info <- sample_info[order(sample_info$Age),]
rownames(sample_info) <- NULL
sample_info$Sample <- factor(sample_info$Sample, levels=sample_info$Sample)
sort_order <- str_remove_all(sample_info$Sample, "[bcf]")
#sort_order <- factor(c("PD43725","PD42036","PD40744","PD42034","PD42563","PD42569","PD40745","PD42564", "PD42565","PD42566","PD42038","PD43727","PD40746","PD42568", "PD43724","PD43726"), levels=c("PD43725","PD42036","PD40744","PD42034","PD42563","PD42569","PD40745","PD42564", "PD42565","PD42566","PD42038","PD43727","PD40746","PD42568", "PD43724","PD43726"))
sort_order <- factor(sort_order, levels=sort_order)


p5 <- plot_spectrum(type_occurrences_late, CT = TRUE, legend = TRUE)
p6 <- plot_spectrum(type_occurrences_late, by = sort_order, CT = TRUE, legend = TRUE)
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/Plots/Mutation_type_plot-maftools.png", res=300, units="in", width = 12, height=7)
grid.arrange(p5, p6, ncol=2, widths=c(1,3))
dev.off()
```

## Plot Early vs Late
```{r}
png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/Plots/Mutation_type_plot-Early_vs_late.png", res=300, units="in", width = 10, height=7)
grid.arrange(p1, p3, p5,  ncol=3, widths=c(3,3,4.7))
dev.off()
```

## Early vs Late with Wilcox.test
```{r}
type_occurrences_early_fract <- type_occurrences_early/rowSums(type_occurrences_early)
type_occurrences_late_fract <- type_occurrences_late/rowSums(type_occurrences_late)


type_occurrences_early_fract[,"Timing"] <- "Early"
type_occurrences_late_fract[,"Timing"] <- "Late"
type_occurrences_early_fract[,"SAMPLE"] <- rownames(type_occurrences_early_fract)
type_occurrences_late_fract[,"SAMPLE"] <- rownames(type_occurrences_late_fract)

type_occurrences_fract <- rbind(type_occurrences_early_fract, type_occurrences_late_fract)
type_occurrences_fract_melt <- melt(type_occurrences_fract, id.vars = c("SAMPLE", "Timing"))
type_occurrences_fract_sum <- summarySE(type_occurrences_fract_melt, measurevar="value", groupvars=c("variable", "Timing"))

for(i in 1:length(type_occurrences_fract[which(type_occurrences_fract$Timing == "Early"),"C>T at CpG"])){
  print(poisson.test(as.integer(type_occurrences_fract[which(type_occurrences_fract$Timing == "Early"),"C>T at CpG"]*100)[[i]],as.integer(type_occurrences_fract[which(type_occurrences_fract$Timing == "Late"),"C>T at CpG"]*100)[[i]],r = 1)$p.value)
}

wilcox_test_pvalues <- data.frame()
for(muttype in colnames(type_occurrences)){
  p <- wilcox.test(as.numeric.Array(type_occurrences_early_fract[,muttype]),as.numeric.Array(type_occurrences_late_fract[,muttype]),paired = T)$p.value
  wilcox_test_pvalues[muttype,"p_value"] <- print(round(p,4))
}

wilcox_test_pvalues["Type"] <- rownames(wilcox_test_pvalues)

to_keep <- c("C>A", "C>G", "C>T at CpG", "C>T other", "T>A", "T>C", "T>G")

type_occurrences_fract_sum <- type_occurrences_fract_sum[which(type_occurrences_fract_sum$variable %in% to_keep),]
wilcox_test_pvalues <-  wilcox_test_pvalues[which(wilcox_test_pvalues$Type %in% to_keep),]
type_occurrences_fract_sum$variable <- factor(type_occurrences_fract_sum$variable, levels=to_keep)

### 95% confidence interval used
g1 <- ggplot(type_occurrences_fract_sum, aes(x=type_occurrences_fract_sum$variable, y=type_occurrences_fract_sum$value, fill=type_occurrences_fract_sum$Timing)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=value-ci, ymax=value+ci),width=.2, position=position_dodge(.9))+scale_fill_brewer(palette = "Dark2", name="")+theme_pubr()+xlab("") + ylab("Relative Contribution") +theme(plot.margin = margin(1,1,1,1, "cm"), axis.text =element_text(size=9))

g2 <- ggplot()+geom_text(data=wilcox_test_pvalues, aes(x=Type,y=0,label = p_value), size=3)+theme_transparent()+theme(plot.margin = margin(-19,0.5,1,2, "cm"))


png("/Users/mn7/volumes/rs30_lustre/Testicular_project/Unmatched_Analysis/Plots/Mutation_type_plot-Early_vs_late_wicoxTestwithCI.png", res=300, units="in", height=5, width=7)
grid.arrange(g1, g2, ncol=1, heights=c(5,0.1))
dev.off()
#..p.format..
#rownames(type_occurrences)
#type_occurrences_early
#type_occurrences_late


```
